---
---
---

```{r}
library(stringr)
# Read lines
ll  = readLines("MergedAuth.log")
ll <- ll[ll != ""]
indices <- grep("^# (.+\\.log)$", ll)
title <- ll[indices]

subll <- ll[indices[4]+1: (length(ll)+1)]

read.log <- function(ll, start, end){
#

  #pattern <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) ((?:[a-zA-Z0-9.-]+(?:\\s*[a-zA-Z0-9.-]+)*(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"
  pattern <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+)\\s+((?:[a-zA-Z0-9.-]+(?:\\s*[a-zA-Z0-9.-]+)*(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"

  # Read line
  subll <- ll[(start+1): (end-1)]
  #testg <- grepl(pattern, auth2_ll)
  # Use str_match() to get the captured groups
  matrix <- str_match(subll, pattern)
  df <- as.data.frame(matrix)
  df<-df[,-1]
  df <- cbind(df, c(ll[start]))
  df[,1] <- as.POSIXct(strptime(df[,1], "%b %d %H:%M:%S"))
  colnames(df) <- c("date-time", "logging host", "app", "PID", "message", "log file")
  df$PID <- as.numeric(df$PID)
  return(df)
}


df1 <- read.log(ll, indices[1], indices[2])
df2 <- read.log(ll, indices[2], indices[3])
df3 <- read.log(ll, indices[3], indices[4])
df4 <- read.log(ll, indices[4], indices[5])
df5 <- read.log(ll, indices[5], length(ll)+1)




df <- rbind(df1, df2, df3, df4, df5)


# Verify
length(ll) - 5 == dim(df1)[1]+dim(df2)[1]+dim(df3)[1]+dim(df4)[1]+dim(df5)[1]

#df4 <- read.log(ll, indices[4], (length(ll)+1))
#df4
#row_indices_with_na <- which(apply(df4, 1, function(x) any(is.na(x))))
#row_indices_with_na
```

```{r}
df <- rbind(df1, df2, df3, df4, df5)
nrow(df)



```

# Data Validation and Exploration

## Verify that the PIDs are all numbers

```{r}
all(sapply(df$PID, function(x) is.numeric(x)))
```

## How many lines are in each log file?

```{r}
dim(df1)[1]
dim(df2)[1]
dim(df3)[1]
dim(df4)[1]
dim(df5)[1]
```

## What are the range of date-times for the messages? for each of the different log files in the combined file? How many

days does each log file span?

```{r}
print("Range of data time for all")
range(df$`date-time`)

print("Range of day in each file")
print(df1$`log file`[1])
range(df1$`date-time`)

print(df2$`log file`[1])
range(df2$`date-time`)

print(df3$`log file`[1])
range(df3$`date-time`)

print(df4$`log file`[1])
range(df4$`date-time`)

print(df5$`log file`[1])
range(df5$`date-time`)

print("days spans in each")
print(df1$`log file`[1])
diff(range(df1$`date-time`))
print(df2$`log file`[1])
diff(range(df2$`date-time`))
print(df3$`log file`[1])
diff(range(df3$`date-time`))
print(df4$`log file`[1])
diff(range(df4$`date-time`))
print(df5$`log file`[1])
diff(range(df5$`date-time`))
```

## Do the application names contain numbers? If so, are they just versions, e.g. ssh2, or is there additional structure to the

numbers?

According to the output, we observe the numbers are separated by periods.

```{r}
a <- unique(df$app)
a[grepl("\\d+", a)]
```

## Is the host value constant/the same for all records in each log file?

Make logging host unique, we found that the host values are not constant.

```{r}
unique(df$`logging host`)
```

## What are the most common apps (daemons/programs) that are logging information on each of the different hosts?

```{r}
print(df1$`log file`[1])
which.max(table(df1$app))
print(df2$`log file`[1])
which.max(table(df2$app))
print(df3$`log file`[1])
which.max(table(df3$app))
print(df4$`log file`[1])
which.max(table(df4$app))
print(df5$`log file`[1])
which.max(table(df5$app))
```

# Logins - valid and invalid

## Find valid/successful logins

### What are the user names and Internet Protocol (IP) addresses of those logins

The valid users are shown below.

```{r}
# Filt all rows realated to login
login_df <- df[grepl("systemd-logind|sshd", df$`app`, ignore.case = TRUE),]
login_df <- login_df[grepl("user", login_df$message, ignore.case = TRUE),]
# make the message unique to see the types
#unique(login_df$message)

```

```{r}
# Filter the invalids
#notlogin <- "server listening|watching system buttons|new seat|removed session|Received signal|Close session"
invalid <- "did not|fatal|invalid|failed|failure|closed|not|warning|disconnect|bad|corrupted|error"
invalid_df <- login_df[grepl(invalid, login_df$message, ignore.case = TRUE),]
valid_df <- login_df[!grepl(invalid, login_df$message, ignore.case = TRUE),]

```

```{r}
user_pattern <- "(?i)(?<=user\\s)([a-z0-9_-]+|\\\\{1,2}\\d+)|(?<=for\\s)(elastic_user_\\d+)"
invalid_user_matrix <- str_match(invalid_df$message, user_pattern)
invalid_user <- invalid_user_matrix[,2]
invalid_user <- unique(invalid_user)
```

```{r}
valid_user_matrix <- str_match(valid_df$message, user_pattern)
valid_user <- valid_user_matrix[,2]
unique(valid_user)
```

## Find the invalid user login/ids

### What were the associated IPs

```{r}
#unique(invalid_df$`logging host`)
ip_pattern <- "(?:from) ([0-9.]+)"
invalid_ip_from_matrix <- str_match(invalid_df$message, ip_pattern)
invalid_ip_from <- unique(invalid_ip_from_matrix)[,2]
invalid_ip_from
```

### Were there multiple invalid user logins from the same IP addresses

Yes

```{r}
library(dplyr)

users <- str_extract(invalid_df$message, user_pattern)

ips <- str_extract(invalid_df$message, "(?<=from)\\s([0-9.]+)")

invalid_user_ip_df <- data.frame(user = users, ip = ips)
invalid_user_ip_df <- na.omit(invalid_user_ip_df)
invalid_user_ip_df
invalid_duplicate_ips <- invalid_user_ip_df %>%
  group_by(ip) %>%
  summarise(num_users = n_distinct(user)) %>%
  filter(num_users > 1)
invalid_duplicate_ips
```

### Were there valid logins from those IP addresses

Yes, there are three of them shown below

```{r}
users <- str_extract(valid_df$message, user_pattern)

ips <- str_extract(valid_df$message, "(?<=from)\\s([0-9.]+)")

valid_user_ip_df <- data.frame(user = users, ip = ips)
valid_user_ip_df <- na.omit(valid_user_ip_df)
valid_user_ip_df
valid_duplicate_ips <- valid_user_ip_df %>%
  group_by(ip) %>%
  summarise(num_users = n_distinct(user)) %>%
  filter(num_users > 1)
valid_duplicate_ips
```

### Are there multiple IPs using the same invalid login?

#### ∗ Are these related IPs, e.g., from the same network/domain?

Yes, and they are all elastic users.

```{r}
duplicate_users <- valid_user_ip_df %>%
  group_by(user) %>%
  summarise(num_ips = n_distinct(ip)) %>%
  filter(num_ips > 1)
duplicate_users
```

### What IP address had too many authentication failures.

188.87.35.25 has the most failures.

```{r}
ip_pattern <- "(?:from) ([0-9.]+)"
invalid_ip_from_matrix <- str_match(invalid_df$message, ip_pattern)
invalid_ip_from <- na.omit(invalid_ip_from_matrix[,2])
which.max(table(invalid_ip_from))
```

# Sudo commands

## What are the executables/programs run via sudo

### -- By what user

The users are all root

```{r}
sudodf <- df[df$`app` == "sudo", ]
sudo_user_pattern <- "(?i)(?<=user\\s)([a-z0-9_-]+|\\\\{1,2}\\d+)|(?i)(?<=user=)([a-z0-9_-]+|\\\\{1,2}\\d+)"
users <- str_extract(sudodf$message, sudo_user_pattern)
userroot <- str_extract(sudodf$message, "(?<=COMMAND)(.*)")
unique(na.omit(userroot))
#sudodf[is.na(users),]
```

### -- What machine are they on?

The machine are the logging host itself, so they are

```{r}
sudodf$`logging host`
```

# TEST CODE CHUNK

```{r, eval=FALSE}
library(stringr)

ll  = readLines("MergedAuth.log")
ll <- ll[ll != ""]
indices <- grep("^# (.+\\.log)$", ll)
ll[indices]
#pattern  <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) ((?:[a-zA-Z0-9.-]+(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"
pattern <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+)\\s+((?:[a-zA-Z0-9.-]+(?:\\s*[a-zA-Z0-9.-]+)*(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"
subll <- ll[(indices[5]+1): (length(ll)+1)]
test <- grepl(pattern, subll)
subll[!test]

#pattern <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) ((?:[a-zA-Z0-9.-]+(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?: (.*)"

#temp <- subll[test][1:10]
#a <- str_match(temp, pattern)
#as.data.frame(a)
log_entry <- c(
  "Jun 15 02:04:59 combo sshd(pam_unix)[20896]: authentication failure; logname= uid=0 euid=0 tty=NODEVssh ruser= rhost=220-135-151-1.hinet-ip.hinet.net  user=root",
  "Dec  1 02:20:05 ip-172-31-27-153 sshd[23842]: Did not receive identification string from 218.75.153.170", 
  "Mar 28 19:00:30 ip-10-77-20-248 systemd-logind[1118]: New session 44 of user ubuntu.",
  "Mar 27 13:08:09 ip-10-77-20-248 systemd: pam_unix(systemd-user:session): session opened for user ubuntu by (uid=0)",
  "Jul  1 09:01:05 calvisitor-10-105-160-95 com.apple.CDScheduler[43]: Thermal pressure state: 1 Memory pressure state: 0",
  "Jul  1 09:29:02 calvisitor-10-105-160-95 sandboxd[129] ([31211]): com.apple.Addres(31211) deny network-outbound /private/var/run/mDNSResponder",
  "Jul  3 17:10:11 calvisitor-10-105-160-184 BezelServices 255.10[94]: ASSERTION FAILED: dvcAddrRef != ((void *)0) -[DriverServices getDeviceAddress:] line: 2789"
  )
#pattern <- "([a-zA-Z]{3} +[0-9]{1,2} [0-9:]+) ([^ ]+) ((?:[a-zA-Z-]+(?:\\([^)]+\\))?)+)\\[(\\d+)\\]: (.*)"
#pattern1 <- "([a-zA-Z]{3} +[0-9]{1,2} [0-9:]+) ([^ ]+) (?:([a-zA-Z0-9\\.\\-]+(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?: (.*)"
#pattern1 <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) ((?:[a-zA-Z0-9.-]+(?:\\([^)]



#test <- grepl(pattern1, log_entry)
#test
#pattern1 <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) (([a-zA-Z0-9.-]+(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"
pattern1 <- "([a-zA-Z]{3} +\\d{1,2} \\d{2}:\\d{2}:\\d{2}) ([^ ]+) ((?:[a-zA-Z0-9.-]+(?:\\s*[a-zA-Z0-9.-]+)*(?:\\([^)]+\\))?)+)(?:\\[(\\d+)\\])?(?:\\s*(?:\\([^)]+\\)))?(?::\\s*)(.*)"




library(stringr)


# Apply str_match() to the log_entry
matches <- str_match(log_entry, pattern1)

# Print the matches matrix
df <- as.data.frame(matches)
df <- df[,-1]
df
```
